Measuring neutron star tidal deformability with Advanced LIGO: a Bayesian analysis of neutron star - black hole binary observations

### 2016-09-05 Paper/paper.tex
improving s4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}\label{s1:techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% #################
% THESE FIGURES ARE FOR THE NEXT SECTION

@@

\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
\textbf{at different SNR values (shown in legend).}\par\medskip
\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
\caption{We show here probability distributions for mass ratio $\eta$ as measured
for the same injection at different SNRs. The intrinsic parameters of the source
are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
{\it ignore} tidal effects.
In each panel: the dashed red line marks the median {\it measured} value
$\eta^\mathrm{Median}$, while the dashed green line show the true value
$\eta^\mathrm{Injected}$. The difference between the true and measured values
is the systematic bias introduced
by ignoring tidal effects in templates. The darker shading shows
the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
which is a direct measure of our statistical uncertainty on $\eta$.
Comparing systematic and statistical errors, we find that:
at $\rho=20$, $\eta$ measurement is dominated by statistical
errors; at $\rho=30$, the two become comparable; and 
for louder signals ($\rho\simeq50$), the systematic errors dominate.
\label{fig:SingleSystemEtaPDFvsSNR}

@@

@@

%%
\textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
\textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
\includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
% \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
% \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
% \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
% \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
We illustrate here two-dimensional joint probability distributions, as
recovered for the intrinsic parameters of a binary with
$q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
$\lambdans=2000$, and $\rho=30$ (same as in
Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
The top row shows the same for combinations of 
non-tidal parameters, i.e. binary masses and spins, while the middle row 
shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
is similar to the middle, with the only difference being that the SNR has been
turned up to $\rho=50$.
\label{fig:SingleSystemLambda2DPDFs}

@@

% set by the detectors).

@@

\label{fig:TT_SNRThresholds_BHspin_BHmass_CI90_0}

@@

% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR20_CI90_0}
% \includegraphics[width=\columnwidth]{plots/TTLambdaErrorCurves_BHspin_BHmass_SNR30_CI90_0}
% \caption{\red{FIGURE FEELS UNFINISHED. POLISH AND SYNCHRONIZE FIGURE WITH CAPTION.}
% \prayush{TRY AND MAKE THIS POINT IN THE FIGURE ABOVE ONLY.}
% In these figures we highlight the regions of BH mass and spin plane 
% where we can constrain the $\lambdans$ of its companion NS within a factor of 
% $1-2$ of its true value. Panels correspond to different signal strength, with 
% $\rho$ increasing from left to right. Within each panel, $\lambdans$ 
% measurement uncertainty contours are shown as functions of BH mass and spin,
% with line-types corresponding to different uncertainty levels, and colors 
% showing the injected $\lambdans$. 
% % We note that this information is, in principle, contained in 
% % Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, which we gather here to better understand
% % the effect of NS's deformability itself on its measurement. 
% As hinted at in Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR}, the 
% measurability of NS tidal deformability improves with all factors that 
% enhances the signature of the NS's disruption on the \textit{detectable} 
% portion of the emitted GW signal (implying, within a frequency band set by the
% detectors).
% \label{fig:TT_LambdaErrorCurves_BHspin_BHmass_CI90_0}

In the previous section, we showed that the effect of deformation of the NS due to
the quadrupolar tidal field of its companion BH becomes discernible in the
GW spectrum only when the following hold:
(a) the BH has a sufficiently small mass, (b) BH spin
is positive aligned, i.e. $\chibh\gtrsim +0.4$, (c) the NS is not very compact,
with $\lambdans\gtrsim 1000$, {\it and} (d) the source is close enough
such that its GW SNR $\gtrsim 30$.
% 
The first two conditions both enhance the tidal distortion of the star and increase
the number of orbits the system goes through at small separation, where the
differences between NSBH and BHBH signals are maximal.
% 
The first three also reduce the onset frequency of the disruption of the NS,
allowing for it to happen earlier in the orbit. 
% 
We expect that these conditions are also the ones which would maximize the
likelihood of {\it measuring} tidal effects in NSBH signals. In this section,
%
%In addition, a SNR of $30$ corresponds to a $1.35+5M_\odot$ binary with 
%$\chibh = 0.4$ at a luminosity distance of $XXX$~Mpc, if optimally aligned
%with LIGO ($XXX$~Mpc if the binary subtends a $45^o$ angle with LIGO's line of
%sight.
%
%These conditions are strong, but not unrealistic. For e.g., a detection rate
%of $20$ NSBHs a year implies $1-2$ of them will have SNRs$\gtrsim 30$, since
%the number of expected events scales as $1/\rho^2$ if we assume that sources
%are distributed uniformly in spacial volume.
%
we turn the question around and ask: under similarly favorable circumstances, can
we gain insights about the internal structure of neutron stars from GW observations.

%
%Ultimately, the extent of tidal moduilations of emitted GWs depends on the
%equation of state of neutron star matter, which becomes an indirect
%measurable in GW astrophysics.
% 
%We therefore turn to the question: If we do observe GWs from disruptive 
%coalescing NSBHs, what information can we extract from them about the internal
%structure of neutron stars?
% 
We begin by calculating the accuracy with which we measure
$\lambdans$ from {\it single} GW observations. We sample the same set of
disruptive NSBH systems as in the previous section, i.e. those with
$q=\{2,3,4,5\}$, $\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$.
For each unique combination of these parameters, we generate separate 
zero-noise injections at SNRs $\rho=\{20,30,50,70\}$, using the LEA+ model. We
perform full Bayesian analyses on each of these injections using LEA+ templates,
and obtain probabilistic estimates for source binary parameters using GW information
alone (including the NS tidal deformability parameter $\lambdans$). Our priors
on component masses and spins remain as in the previous section, while $\lambdans$
is sampled uniformly from $[0,4000]$, and mass-ratio is additionally restricted to
$2\leq q\leq 6$.

As an illustration, we show the recovered
probability distribution for $\lambdans$ for three specific configurations in 
Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix $q = m_1/m_2 =
5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while varying $\lambdans$
over $\{1000, 1500, 2000\}$ between the three panels. The SNR is fixed at $\rho=50$.
The true value of $\lambdans$ is given the green (dashed) vertical line,
while the median recovered value is marked by a red (dashed) line. The darker
shaded region marks the $90\%$ credible interval on $\lambdans$. One thing
to note in this figure is that the measurement uncertainty on $\lambdans$, i.e.
the full width of its $90\%$ credible interval, grows slower than $\lambdans$
as $\lambdans$ is turned up (while the fractional uncertainty
\begin{equation}
\delta\lambdans^{90\%}:= (\Delta\lambdans)^{90\%}/\lambdans
\end{equation}
decreases with increasing $\lambdans$).
%
Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
mass, spin, and tidal parameter measurements. We keep the binary parameters
as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
$\rho=30$ (instead of $50$). In the top panel, we show the correlation between
non-tidal parameters, as manifested in their recovered joint probability
distributions.
In the middle row, we show the same between $\lambdans$ and non-tidal 
parameters, i.e. chirp mass (left panel), mass ratio $\eta$ (middle panel), and
BH spin $\chibh$ (right panel). The same is shown in the row below, with the
difference that the SNR has been turned up to $\rho=50$, which brings out the
intrinsic ambiguity contours more clearly.
%
We find that the measurement of $\lambdans$ is weakly degenerate with
other parameters, and at realistic SNRs it would improve by a few tens of 
percent if we knew non-tidal parameters to better accuracy. The predominant
factor that would enhance the measurement accuracy for $\lambdans$ is the
signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
measurement to be limited by its degeneracy with non-tidal parameters at 
a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.

In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the main results
of this section. In each panel, as a function of black hole mass and spin,
we show the measured $90\%$ credible intervals for $\lambdans$, i.e.
$(\Delta\lambdans)^{90\%}$. These correspond to the full width of the dark
shaded region in the illustrative figure,
Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. The effect of signal strength,
i.e. SNR, can be seen as we go from left to right in each row. The
effect of the tidal deformability of the NS $\lambdans$ on its own measurability
can be seen by comparing panels within each column. The NS becomes more 
deformable as we go from top to bottom. In addition, contours of
$\dlambda=\{50\%,75\%,100\%,150\%,200\%\}$ are also shown in all panels.
%
A clear pattern emerges in the left-most column, which corresponds to
$\rho=20$. We find that at this signal strength, our measurement of 
$\lambdans$ is dominated by the width of our prior on $\lambdans$. In other
words, our measured credible interval is restricted by the
allowed range of $\lambdans$. Increasing the signal strength to
$\rho=30$ gives somewhat similar results, as can be seen from the second-from-left
column of the figure. We do, however, find more encouraging results
for high BH spins, in which case we can put $\pm 100\%$ to $\pm 75\%$
bounds on $\lambdans$~\footnote{The symmetric error-bars of
$\pm\mathrm{X}\%$ correspond to $\dlambda = 2\mathrm{X}\%$.}.
In addition to large BH spins, this level of
precision will also require rather deformable neutron stars.
%
At higher SNRs, $\rho\gtrsim 50$, we can put stronger constraints
on $\lambdans$. With a {\it single} observation of a $q=4$ binary with
$\chibh\geq 0.6$ and $\rho = 50$~\footnote{For an optimally oriented
source with $q=4, \mns=1.35M_\odot, \chibh=0.6$, an SNR of 
$\rho = 50$ corresponds to a luminosity distance of $\approx 113$Mpc.},
we would be able to estimate 
$\lambdans$ to within $\pm 40\%$ of its true value 
(which is equivalent to measuring the ratio of its radius to mass
with an uncertainty of about $\pm 10\%$).

@@

% #################
\begin{figure*}
\centering    
\textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
\includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
\includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
\caption{%
These panels show the accumulation of information from a population whose
$\lambdans=800$.
{\it Left:} Posterior probability distributions for $\lambdans$ (colored 
curves), and associated $90\%$ credible intervals (grey vertical lines),
shown for different number of accumulated observations $N$. Distributions are
normalized to unit area.
{\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
associated $90\%$ credible intervals (as the vertical extent of filled
region), shown as a function of number of observations $N$. Solid horizontal
line indicates the true value of $\lambdans$. Dashed and dotted horizontal
lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
bounds.
}
\label{fig:TT_Lambda_vs_N_L800_CI90_0}
\end{figure*}
%
% #################

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Techniques}\label{s1:techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% #################
% THESE FIGURES ARE FOR THE NEXT SECTION
\begin{figure*}
\centering 
\textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
\textbf{at different SNR values (shown in legend).}\par\medskip
\includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
\caption{We show here probability distributions for mass ratio $\eta$ as measured
for the same injection at different SNRs. The intrinsic parameters of the source
are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
{\it ignore} tidal effects.
% 
In each panel: the dashed red line marks the median {\it measured} value
$\eta^\mathrm{Median}$, while the dashed green line show the true value
$\eta^\mathrm{Injected}$. The difference between the true and measured values
is the systematic bias introduced
by ignoring tidal effects in templates. The darker shading shows
the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
which is a direct measure of our statistical uncertainty on $\eta$.
% 
Comparing systematic and statistical errors, we find that:
at $\rho=20$, $\eta$ measurement is dominated by statistical
errors; at $\rho=30$, the two become comparable; and 
for louder signals ($\rho\simeq50$), the systematic errors dominate.
}
\label{fig:SingleSystemEtaPDFvsSNR}
\end{figure*}

@@

% \textbf{Illustrative posterior probability distributions for mass-ratio $\eta$,}\par
% \textbf{at different SNR values (shown in legend).}\par\medskip
% \includegraphics[width=1.8\columnwidth]{plots/SingleSystemEta_q4_0_mc2_25_chi0_50}
% \caption{We show here probability distributions for mass ratio $\eta$ as measured
% for the same injection at different SNRs. The intrinsic parameters of the source
% are: $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and $\lambdans=2000$;
% and the signal is injected at SNRs $\rho=\{20,30,50\}$ (left to right). The templates
% {\it ignore} tidal effects.
% In each panel: the dashed red line marks the median {\it measured} value
% $\eta^\mathrm{Median}$, while the dashed green line show the true value
% $\eta^\mathrm{Injected}$. The difference between the true and measured values
% is the systematic bias introduced
% by ignoring tidal effects in templates. The darker shading shows
% the recovered $90\%$ credible interval for $\eta$, $(\Delta\eta)^{90\%}$, 
% which is a direct measure of our statistical uncertainty on $\eta$.
% at $\rho=20$, $\eta$ measurement is dominated by statistical
% \label{fig:SingleSystemEtaPDFvsSNR}

@@

% ##############################################################

@@

% %% FIXME FIXME
% \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
% \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
% \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
% % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
% % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
% % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
% % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
% % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
% % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
% % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
% We illustrate here two-dimensional joint probability distributions, as
% recovered for the intrinsic parameters of a binary with
% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
% $\lambdans=2000$, and $\rho=30$ (same as in
% Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
% The top row shows the same for combinations of 
% non-tidal parameters, i.e. binary masses and spins, while the middle row 
% shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
% is similar to the middle, with the only difference being that the SNR has been
% turned up to $\rho=50$.
% \label{fig:SingleSystemLambda2DPDFs}
% \textbf{Statistical uncertainty in $\lambdans$ measurement}\par\medskip
% \includegraphics[trim={2cm 0 0 0},width=2.2\columnwidth]{plots/TTLambdaRawCIWidths90_0_Lambda_SNR}
% % \includegraphics[trim={3cm 0 0 0},width=2.\columnwidth]{plots/TTLambdaCIWidths90_0_Lambda_SNR.pdf}
% \caption{Here we show the statistical uncertainty in the measurement of
% $\lambdans$, as a percentage of the injected/true value. In each panel, the
% same is shown as a function of the BH mass and spin, keeping $\lambdans$ and
% injection's SNR $\rho$ fixed (noted in the panel). Each row contains panels
% with the same value of $\lambdans$, with $\rho$ increasing from left to right.
% Each column contains panels with the same value of $\rho$, with $\lambdans$ 
% increasing from top to bottom.
% Contours demarcate regions where we can constrain the
% $\lambdans$ parameter well (within a factor of two of the injected value).
% We note that, as expected, the measurement accuracy for $\lambdans$ improves with (i) increasing
% SNR, (ii) decreasing BH mass, (iii) increasing BH spin, and 
% (iv) increasing $\lambdans$, i.e. the tidal deformability of the neutron star.
% \label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
% \textbf{Illustrative posterior probability distributions for NS tidal
% deformability $\lambdans$,}\par
% \textbf{for signals with different $\lambdans$ values; $\lambdans = 1000,1500,2000,$
% from left to right.}\par\medskip
% % \includegraphics[width=1.8\columnwidth]{plots/SingleSystem_q4_0_mc2_25_chi0_50}
% \includegraphics[width=1.9\columnwidth]{plots/SingleSystemLambdaVary_q4_0_mc2_25_chi0_50_snr50}
% \caption{We illustrate here probability distributions recovered for the NS tidal
% deformability parameter $\lambdans$ from three GW injections, with parameters:
% $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$, and 
% $\lambdans=\{1000,1500,2000\}$ from left to right. The injection SNR is fixed at
% $\rho=50$. The templates {\it include} tidal effects, with a prior $0\leq\lambdans\leq 4000$.
% % These figures show that this approximation holds up to (SNR)
% % $\rho\simeq 30-50$.
% In each panel- the dashed red line marks the median {\it measured} value for
% $\lambdans$, and the dashed green line marks its {\it true} value.
% The darker shading shows the $90\%$ credible interval, whose width
% $(\Delta\lambdans)^{90\%}$ is a direct measure of our statistical uncertainty.
% By comparing the measurement uncertainty for these three injections, we see
% that $(\Delta\lambdans)^{90\%}$ grows very slowly with $\lambdans$. Therefore,
% the fractional measurement error - $(\Delta\lambdans)^{90\%}/\lambdans$ -
% decreases monotonically as $\lambdans$ increases (with signal strength fixed).
% % Comparing systematic and statistical errors, we find that:
% % at $\rho=20$, $\lambdans$ measurement is dominated by statistical
% % errors; at $\rho=30$, the two become comparable; and 
% % for louder signals ($\rho\simeq50$), the systematic errors dominate.
% \label{fig:SingleSystemLambdaPDFvsSNR}
% % %% FIXME FIXME
% % \begin{figure*}
% % \centering
% % \textbf{Illustrative posterior probability distributions for NSBH parameters,}\par
% % \textbf{for signals at different SNR values; $\rho = 20$ (left) and $50$ (right).}\par\medskip
% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr20}%\\
% % \includegraphics[width=1.05\columnwidth,trim=2cm 0 0 0]{plots/AllParamsMcEtPDF1D2D_q4_mc2_25_chi0_50_snr50}%\\
% % % \includegraphics[width=0.5\columnwidth]{plots/McEta2D_q4_mc2_25_chi0_50_snr30}
% % % \includegraphics[width=0.5\columnwidth]{plots/EtaChiBH2D_q4_mc2_25_chi0_50_snr30}
% % % \includegraphics[width=0.5\columnwidth]{plots/McChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr30}
% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr30}
% % % \includegraphics[width=0.5\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr30}\\
% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaMc2D_q4_mc2_25_chi0_50_snr50}
% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaEta2D_q4_mc2_25_chi0_50_snr50}
% % % \includegraphics[width=0.47\columnwidth]{plots/LambdaChiBH2D_q4_mc2_25_chi0_50_snr50}
% % \caption{%
% % We illustrate here two-dimensional joint probability distributions, as
% % recovered for the intrinsic parameters of a binary with
% % $q = \mbh/\mns = 5.4M_\odot/1.35M_\odot = 4$, $\chibh=+0.5$,
% % $\lambdans=2000$, and $\rho=30$ (same as in
% % Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}).
% % The top row shows the same for combinations of 
% % non-tidal parameters, i.e. binary masses and spins, while the middle row 
% % shows combinations of $\lambdans$ with non-tidal parameters. The bottom row
% % is similar to the middle, with the only difference being that the SNR has been
% % turned up to $\rho=50$.
% % }
% % \label{fig:SingleSystemLambda2DPDFs}
% % \end{figure*}

@@

% set by the detectorTT_SNRThresholds_BHspin_BHmass_CI90_0s).

@@

\label{fig:}

@@

In the previous section, we showed that the effects of the deformation of NSs
due to the quadrupolar tidal field of their companion BHs become discernible in
the GW spectrum under certain favorable conditions, including (a) BH mass is
sufficiently small, (b) BH spin is positive aligned, i.e. $\chibh\gtrsim +0.4$,
(c) the NS is not very compact, with $\lambdans\gtrsim 1000$, and (d) the
source location and orientation are such that its GW SNR $\gtrsim 30$.
% 
Both condition (a) and (b) enhance the tidal distortion of the star and increase
the number of orbits the system goes through at small separation, where the
differences between NSBH and BBH signals are maximal.
% 
Conditions (a)-(c) also reduce the onset frequency of the disruption of the NS,
allowing for it to happen earlier in the orbit. 
% 
We expect that these conditions are also the ones which would maximize the
likelihood of {\it measuring} tidal effects in NSBH signals. In this section,
%
%In addition, a SNR of $30$ corresponds to a $1.35+5M_\odot$ binary with 
%$\chibh = 0.4$ at a luminosity distance of $XXX$~Mpc, if optimally aligned
%with LIGO ($XXX$~Mpc if the binary subtends a $45^o$ angle with LIGO's line of
%sight.
%
%These conditions are strong, but not unrealistic. For e.g., a detection rate
%of $20$ NSBHs a year implies $1-2$ of them will have SNRs$\gtrsim 30$, since
%the number of expected events scales as $1/\rho^2$ if we assume that sources
%are distributed uniformly in spacial volume.
%
we turn the question around and ask: under similarly favorable circumstances, can
we gain insights about the internal structure of neutron stars from GW observations.

In this section, we calculate the accuracy with which we measure $\lambdans$ from
{\it single} GW observations. We sample the same set of disruptive NSBH mergers
as in the previous section, i.e. those with $q=\{2,3,4,5\}$,
$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000, 1500, 2000\}$;
fixing the NS mass $\mns=1.35 M_\odot$ and $\chins=0$. For each unique
combination of these parameters, we inject LEA+ signals into zero noise and
perform a fully Bayesian parameter estimation analysis of each with LEA+
templates. Our priors on component masses and spins remain as in the previous
section, with mass-ratio additionally restricted to $2\leq q\leq 6$, and
$\lambdans$ sampled uniformly from $[0,4000]$.
% 
% 
%
As an illustration, we show the recovered probability distribution for
$\lambdans$ for three specific configurations in 
Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. We fix
$q = \mbh /\mns = 5.4M_\odot/1.35M_\odot = 4$, with $\chibh=+0.5$, while
varying $\lambdans$ over $\{1000, 1500, 2000\}$ between the three panels.
The SNR is fixed at $\rho=50$. Both the true and measured values of
$\lambdans$ are marked. The darker shaded region marks the $90\%$ credible
interval on $\lambdans$. We note that $\lambdans$ is estimated to within
$\pm 2000$ of its true value at this SNR. Another interesting thing to note 
is that while $(\Delta\lambdans)^{90\%}$ slowly grows with $\lambdans$, the
fractional uncertainty
\begin{equation}
\delta\lambdans^{90\%}:= (\Delta\lambdans)^{90\%}/\lambdans
\end{equation}
decreases instead.
%
\textcolor{blue}{%
Further illustrations of the measurements of mass, spin and tidal parameters,
as well as their correlations, are shown in Appendix~\ref{as1:illustrations}.
}
% 
\textcolor{gray}{%
Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
mass, spin, and tidal parameter measurements. We keep the binary parameters
as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
set $\rho=30$ (left panel) or $\rho=50$ (right panel).
%
We find that the measurement of $\lambdans$ is weakly degenerate with
other parameters, and at realistic SNRs it would improve by a few tens of 
percent if we knew non-tidal parameters to better accuracy. The predominant
factor that would enhance the measurement accuracy for $\lambdans$ is the
signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
measurement to be limited by its degeneracy with non-tidal parameters at 
a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
}

In Fig.~\ref{fig:TT_LambdaCIWidths90_0_Lambda_SNR} we show the main results of
this section. In each panel, as a function of black hole mass and spin, we show
the measured $90\%$ credible interval widths $(\Delta\lambdans)^{90\%}$. These
correspond to the full width of the dark shaded regions in the illustrative
Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}. The effect of increasing signal
strength can be seen as we go from left to right in each row. The effect of the
NS tidal deformability parameter $\lambdans$ on its own measurability can be
seen by comparing panels within each column, with the NS becoming more
deformable from top to bottom. 
%
A uniform pattern emerges in the left-most column, which corresponds to $\rho=20$.
We find that at this signal strength, our measurement of $\lambdans$ is
dominated by the width of our prior on it. The $90\%$ credible intervals span
the entire allowed range for $\lambdans$, making a reasonable estimation of
$\lambdans$ at $\rho\simeq20$ difficult.
% 
Increasing the signal strength to $\rho=30$ gives marginally better results,
bringing down the statistical uncertainties to within $\pm 75-100\%$ of the
true $\lambdans$ value~\footnote{The symmetric error-bars of $\pm\mathrm{X}\%$
correspond to $\dlambda = 2\mathrm{X}\%$.}.
%
It is not until we reach an SNR as high as $\rho\gtrsim 50$, can we put
meaningful (i.e. $\mathcal{O}(10\%)$) constraints on $\lambdans$. For e.g.,
with a {\it single} observation of a $q=4$ binary with $\chibh\geq 0.6$ and
$\rho = 50$~\footnote{For an optimally oriented source with
$q=4, \mns=1.35M_\odot, \chibh=0.6$, an SNR of $\rho = 50$ corresponds to
a luminosity distance of $\approx 113$Mpc.}, we would be able to estimate 
$\lambdans$ to within $\pm 40\%$ of its true value (which is equivalent to
measuring the ratio of NS radius to mass with an uncertainty of about
$\pm 10\%$).
% 
The results presented so far are quite in agreement with Sec.~\ref{s1:PEwithnoNS}.

@@

% #################
% \begin{figure*}
% \centering    
% \textbf{Improvement in $\lambdans$ measurement accuracy with multiple observations}\par\medskip
% \includegraphics[width=1.05\columnwidth,trim=1cm 0 0 0]{plots/pdfLambda_vs_N_L800.pdf}
% \includegraphics[width=1.\columnwidth,trim=0 0 1cm 0]{plots/FillBetweenErrorBarsLambda_vs_N_L800.pdf}
% \caption{%
% These panels show the accumulation of information from a population whose
% $\lambdans=800$.
% {\it Left:} Posterior probability distributions for $\lambdans$ (colored 
% curves), and associated $90\%$ credible intervals (grey vertical lines),
% shown for different number of accumulated observations $N$. Distributions are
% normalized to unit area.
% {\it Right:} Measured median value of $\lambdans$ (as solid circles) and the 
% associated $90\%$ credible intervals (as the vertical extent of filled
% region), shown as a function of number of observations $N$. Solid horizontal
% line indicates the true value of $\lambdans$. Dashed and dotted horizontal
% lines (a pair for each line-style) demarcate $\pm 25\%$ and $\pm 50\%$ error
% bounds.
% }
% \label{fig:TT_Lambda_vs_N_L800_CI90_0}
% \end{figure*}
% %
% #################

@@

Next, in Fig.~\ref{fig:SingleSystemLambda2DPDFs} we show the correlation of
mass, spin, and tidal parameter measurements. We keep the binary parameters
as in Fig.~\ref{fig:SingleSystemLambdaPDFvsSNR}, with $\lambdans=2000$, and
set $\rho=30$ (left panel) or $\rho=50$ (right panel).
%
We find that the measurement of $\lambdans$ is weakly degenerate with
other parameters, and at realistic SNRs it would improve by a few tens of 
percent if we knew non-tidal parameters to better accuracy. The predominant
factor that would enhance the measurement accuracy for $\lambdans$ is the
signal strength, and only when $\rho\gtrsim 50$ do we expect $\lambdans$ 
measurement to be limited by its degeneracy with non-tidal parameters at 
a factor of few level, as was reported by Ref.~\cite{Lackey:2013axa}.
